\documentclass[10pt]{beamer}

\usetheme{m}

\usepackage{booktabs}
\usepackage{fontawesome}
\usepackage{tikz}
\usetikzlibrary{arrows,positioning}

\title{Bloc: Binary Large Objects with Concurrency}
\subtitle{}
\date{\today}
\author{Anshu Avinash \\
  Thesis supervisor: Piyush Kurur}
\institute{Department of CSE, IIT Kanpur}

\begin{document}

\maketitle

\begin{frame}
  \frametitle{What is a blob?}

  BLOB stands for ``\textbf{B}inary \textbf{L}arge \textbf{OB}ject''.
  \begin{center}
    \begin{tikzpicture}[>=latex]
      \node (picture) {\Huge \faPicture};
      \node [below =1cm of picture] (film) {\Huge \faFilm};
      \node [below =1cm of film] (music) {\Huge \faMusic};
      \node [below right =1cm and 2cm of picture] (database) {\Huge \faDatabase};

      \path[draw,->] (picture) edge (database)
      (film) edge (database)
      (music) edge (database)
      ;
    \end{tikzpicture}
  \end{center}

\end{frame}

\begin{frame}[allowframebreaks]
  \frametitle{How to store blobs?}

  \textbf {Store the entire content of the blob in the database.}
  \begin{itemize}
    \item PostgreSQL breaks large objects into ``chunks'' and these chunks are
      stored in rows in the database.
    \item MongoDB also divides large objects using the GridFS specification and stores
      them in the ``chunks'' collection.
  \end{itemize}

  \framebreak

  \textbf {Store the blob in a file and store the file name in the database.}
  \begin{itemize}
    \item Number and size of blobs are now limited only by the file system.
    \item Requires to develop an interface to keep track of all the blobs of a database.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Here comes Bloc!}

  We provide a library which will keep track of all the blobs stored in a database. \\

\end{frame}

\begin{frame}
  \frametitle{Design}
  \begin{itemize}[<+->]
    \item All the blobs of a database are stored inside a directory which we also call a \alert{BlobStore}.
    \item Each blob is a stored as a different file inside the BlobStore.
  \end{itemize}
\end{frame}

\end{document}
